---
title: "Getting Started with bridgr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with bridgr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

library(bridgr)
```

# Overview

`bridgr` is a mixed-frequency forecasting package for bridge models, MIDAS-style
regressions, and intermediate specifications that estimate within-period
weights from the data.

The core workflow is always the same:

1. Provide one lower-frequency target series and one or more
   higher-frequency indicators.
2. Decide how missing indicator observations should be handled with
   `indic_predict`.
3. Decide how indicators should be aligned to the target frequency with
   `indic_aggregators`.
4. Fit the target equation with `mf_model()`, inspect it with `summary()`, and
   produce target-period forecasts with `forecast()`.

This vignette walks through that workflow with the package's built-in Swiss
GDP and barometer data.

# Example Data

```{r data}
gdp_growth <- suppressMessages(tsbox::ts_na_omit(tsbox::ts_pc(gdp)))

head(gdp_growth)
head(baro)
```

`gdp_growth` is quarterly, while `baro` is monthly. `mf_model()` recognizes the
frequency mismatch automatically and aligns the indicator to the target
frequency before fitting the target equation.

# A Basic Bridge Model

```{r basic-model}
bridge_model <- mf_model(
  target = gdp_growth,
  indic = baro,
  indic_predict = "auto.arima",
  indic_aggregators = "mean",
  indic_lags = 1,
  target_lags = 1,
  h = 2
)

forecast(bridge_model)
```

The default mean aggregator is the classic bridge-model setup: each monthly
block is completed first, then averaged to the quarterly frequency before the
target equation is estimated.

The fitted object stores the aligned data that went into estimation and the
future target-period regressor path used for forecasting.

```{r aligned-data}
tail(bridge_model$estimation_set)
bridge_model$forecast_set
```

# Standardized Output

`summary()` and `forecast()` use a stable package-specific layout. The base
output is the same across bridge, mixed-frequency, and direct-alignment
specifications. Additional details, such as optimization summaries or
uncertainty settings, are appended only when they are relevant.

```{r model-summary}
summary(bridge_model)
```

# Forecast Visualization

The package also provides a built-in plotting method for fitted bridge models.
With `type = "forecast"`, it shows the observed target history together with
the forecast path generated by the model.

```{r forecast-plot}
plot(bridge_model, type = "forecast")
```

# Direct Alignment

If you set `indic_predict = "direct"`, `bridgr` switches from indicator
forecasting to direct alignment based only on observed complete
high-frequency blocks. In that case, the latest complete blocks are assigned
backward to target periods instead of being forecast forward first, and they
are averaged within each target period.

```{r direct-model}
direct_model <- mf_model(
  target = gdp_growth,
  indic = baro,
  indic_predict = "direct",
  h = 1
)

forecast(direct_model)
```

This is particularly useful at the ragged edge when you want to work only with
observed high-frequency information and avoid a separate indicator forecasting
step.

# Optional Uncertainty Output

By default, `bridgr` returns point forecasts only. If you want uncertainty
output, request it at estimation time with `se = TRUE` and, if needed,
custom simulation or full-system bootstrap controls through `bootstrap`.

```{r uncertainty-example}
uncertainty_model <- mf_model(
  target = gdp_growth,
  indic = baro,
  indic_predict = "auto.arima",
  indic_aggregators = "mean",
  target_lags = 1,
  h = 4,
  se = TRUE,
  bootstrap = list(N = 40, block_length = NULL)
)

forecast(uncertainty_model)
summary(uncertainty_model)
plot(uncertainty_model, type = "forecast")
```

The uncertainty implementation uses HAC standard errors for the linear bridge
equation, or Delta-HAC standard errors when parametric aggregation weights are
estimated jointly. By default, prediction intervals are simulated from
resampled centered target-equation residuals. If you also set
`full_system_bootstrap = TRUE`, `bridgr` instead uses a full-system
target-period block bootstrap for both coefficient standard errors and
prediction intervals, controlled through `bootstrap = list(N = ...,
block_length = ...)`.

# Where to Go Next

The vignette
`vignette("mixed-frequency-modeling", package = "bridgr")` compares the main
aggregation strategies and shows how `bridgr` moves from classic bridge models
to unrestricted and parametric MIDAS-style specifications.

The vignette
`vignette("ragged-edge-nowcasting", package = "bridgr")` focuses on
`indic_predict` and the different ways to handle incomplete high-frequency
data at the forecast origin.

The vignette
`vignette("uncertainty-and-scenarios", package = "bridgr")` shows how to work
with HAC / Delta-HAC coefficient uncertainty, residual-resampling prediction
intervals, the optional full-system bootstrap, and scenario forecasts based on
custom future regressor paths.
