---
title: "Getting Started with bridgr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with bridgr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
set.seed(42)
library(bridgr)
```

# Overview

`bridgr` covers classic bridge equations, MIDAS-style mixed-frequency
regressions, and intermediate specifications that estimate within-period
weights from the data, all behind a single workflow for **bridging
high-frequency indicators to a lower-frequency target**.

The core workflow is always the same:

1. Provide one lower-frequency target series and one or more
   higher-frequency indicators.
2. Decide how missing indicator observations should be handled with
   `indic_predict`.
3. Decide how indicators should be aligned to the target frequency with
   `indic_aggregators`.
4. Fit the target equation with `mf_model()`, inspect it with `summary()`,
   and produce target-period forecasts with `forecast()`.

This vignette walks through that workflow with the package's built-in Swiss
GDP and indicator data.

# Example Data

```{r data}
gdp_growth <- suppressMessages(tsbox::ts_na_omit(tsbox::ts_pc(gdp)))

head(gdp_growth)
head(baro)
```

`gdp_growth` is quarterly, while `baro` is monthly. `mf_model()` recognizes
the frequency mismatch automatically and aligns the indicator to the target
frequency before fitting the target equation.

## Stationarity expectations

`bridgr` assumes that the series you submit are already on a scale that is
suitable for linear regression — typically growth rates, differences, or
similarly stabilized transformations. The package does not automatically
re-transform inputs, but you can enable a lightweight pre-fit diagnostic
through `stationarity = "warn"`. The example below shows the warning on a
level GDP series and confirms that it disappears after differencing to
growth rates.

```{r stationarity, warning = TRUE}
gdp_level_recent <- dplyr::slice_tail(gdp, n = 40)
baro_recent <- dplyr::slice_tail(baro, n = 120)

invisible(tryCatch(
  mf_model(
    target = gdp_level_recent,
    indic = baro_recent,
    indic_predict = "last",
    indic_aggregators = "mean",
    h = 1,
    stationarity = "warn"
  ),
  warning = function(w) message("Warning captured: ", conditionMessage(w))
))

stationary_model <- mf_model(
  target = gdp_growth,
  indic = baro,
  indic_predict = "last",
  indic_aggregators = "mean",
  h = 1,
  stationarity = "warn"
)
```

# A Basic Bridge Model

```{r basic-model}
bridge_model <- mf_model(
  target = gdp_growth,
  indic = baro,
  indic_predict = "auto.arima",
  indic_aggregators = "mean",
  indic_lags = 1,
  target_lags = 1,
  h = 2
)

forecast(bridge_model)
```

The default `"mean"` aggregator is the classic bridge-model setup: each
monthly block is completed first, then averaged to the quarterly frequency
before the target equation is estimated.

The fitted object stores the aligned data that went into estimation and the
future target-period regressor path used for forecasting.

```{r aligned-data}
tail(bridge_model$estimation_set)
bridge_model$forecast_set
```

## Lags in the target equation

`target_lags = p` adds `p` autoregressive lags of the target to the
right-hand side of the regression. `indic_lags = q` adds `q` lags of *each
aggregated indicator* (after frequency alignment). Both default to `0`. In
the example above the target equation is

$$
\Delta y_t = \alpha
  + \rho \, \Delta y_{t-1}
  + \beta_0 \, \bar{x}_t
  + \beta_1 \, \bar{x}_{t-1}
  + \varepsilon_t,
$$

where $\bar{x}_t$ is the within-quarter mean of the monthly indicator.
Forecasts beyond the first horizon are produced recursively, with simulated
target lags taking the place of observed ones once the in-sample history is
exhausted.

# Standardized Output

`summary()` and `forecast()` use a stable package-specific layout. The base
output is the same across bridge, mixed-frequency, and direct-alignment
specifications. Additional details, such as optimization summaries or
uncertainty settings, are appended only when they are relevant.

```{r model-summary}
summary(bridge_model)
```

# Forecast Visualization

The package also provides a built-in plotting method for fitted
mixed-frequency models. With `type = "forecast"`, it shows the observed
target history together with the forecast path generated by the model.

```{r forecast-plot}
plot(bridge_model, type = "forecast")
```

# Multiple Indicators

`mf_model()` accepts any number of indicators in a single long-format
table. Indicators may share a frequency or come from different frequencies,
in which case each one is aligned to the target separately. The example
below pairs the level of the KOF barometer with its year-over-year change
as two monthly indicators.

```{r multi-indicator}
baro_yoy <- baro |>
  dplyr::arrange(time) |>
  dplyr::mutate(values = values - dplyr::lag(values, 12)) |>
  dplyr::filter(!is.na(values))

indic_multi <- dplyr::bind_rows(
  dplyr::mutate(baro, id = "baro_level"),
  dplyr::mutate(baro_yoy, id = "baro_yoy")
) |>
  dplyr::select(id, time, values)

multi_model <- mf_model(
  target = gdp_growth,
  indic = indic_multi,
  indic_predict = c("last", "last"),
  indic_aggregators = c("mean", "mean"),
  target_lags = 1,
  h = 1
)

multi_model$regressor_names
forecast(multi_model)
```

Both indicators are aggregated independently and enter the final target
equation as separate regressors.

# Direct Alignment

If you set `indic_predict = "direct"`, `bridgr` switches from indicator
forecasting to direct alignment based only on observed complete
high-frequency blocks. In that case, the latest complete blocks are assigned
backward to target periods instead of being forecast forward first, and they
are averaged within each target period.

```{r direct-model}
direct_model <- mf_model(
  target = gdp_growth,
  indic = baro,
  indic_predict = "direct",
  h = 1
)

forecast(direct_model)
```

This is particularly useful at the ragged edge when you want to work only
with observed high-frequency information and avoid a separate indicator
forecasting step.

# Non-Standard Calendars

By default the package uses a regular frequency ladder
(`second` → `minute` → `hour` → `day` → `week` → `month` → `quarter` →
`year`) with conventional conversion factors such as 60 seconds per minute,
24 hours per day, and 3 months per quarter. If your data follow a
non-standard calendar — for example a 5-day business week or 50 working
weeks per year — pass a named numeric vector to `frequency_conversions` to
override the defaults. See `?mf_model` for the full set of recognized
names.

# Optional Uncertainty Output

By default, `bridgr` returns point forecasts only. If you want uncertainty
output, request it at estimation time with `se = TRUE` and, if needed,
custom simulation or full-system bootstrap controls through `bootstrap`.

```{r uncertainty-example}
uncertainty_model <- mf_model(
  target = gdp_growth,
  indic = baro,
  indic_predict = "auto.arima",
  indic_aggregators = "mean",
  target_lags = 1,
  h = 4,
  se = TRUE,
  bootstrap = list(N = 40, block_length = NULL)
)

forecast(uncertainty_model)
summary(uncertainty_model)
plot(uncertainty_model, type = "forecast")
```

The uncertainty implementation uses HAC standard errors for the linear
target equation, or Delta-HAC standard errors when parametric aggregation
weights are estimated jointly. By default, prediction intervals are
simulated from resampled centered target-equation residuals. If you also
set `full_system_bootstrap = TRUE`, `bridgr` instead uses a full-system
target-period block bootstrap for both coefficient standard errors and
prediction intervals, controlled through
`bootstrap = list(N = ..., block_length = ...)`.

# Where to Go Next

The vignette
`vignette("mixed-frequency-modeling", package = "bridgr")` compares the
main aggregation strategies and shows how `bridgr` moves from classic
bridge models to unrestricted and parametric MIDAS-style specifications.

The vignette
`vignette("ragged-edge-nowcasting", package = "bridgr")` focuses on
`indic_predict` and the different ways to handle incomplete high-frequency
data at the forecast origin.

The vignette
`vignette("uncertainty-and-scenarios", package = "bridgr")` shows how to
work with HAC / Delta-HAC coefficient uncertainty, residual-resampling
prediction intervals, the optional full-system bootstrap, and scenario
forecasts based on custom future regressor paths.
