---
title: "Getting Started with bridgr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with bridgr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

library(bridgr)
```

# Overview

`bridgr` is a mixed-frequency forecasting package for bridge models, MIDAS-style
regressions, and intermediate specifications that estimate within-period
weights from the data.

The core workflow is always the same:

1. Provide one lower-frequency target series and one or more
   higher-frequency indicators.
2. Decide how missing indicator observations should be handled with
   `indic_predict`.
3. Decide how indicators should be aligned to the target frequency with
   `indic_aggregators`.
4. Fit the target equation with `bridge()`, inspect it with `summary()`, and
   produce target-period forecasts with `forecast()`.

This vignette walks through that workflow with the package's built-in Swiss
GDP and barometer data.

# Example Data

```{r data}
gdp_growth <- suppressMessages(tsbox::ts_na_omit(tsbox::ts_pc(gdp)))

head(gdp_growth)
head(baro)
```

`gdp_growth` is quarterly, while `baro` is monthly. `bridge()` recognizes the
frequency mismatch automatically and aligns the indicator to the target
frequency before fitting the target equation.

# A Basic Bridge Model

```{r basic-model}
bridge_model <- bridge(
  target = gdp_growth,
  indic = baro,
  indic_predict = "auto.arima",
  indic_aggregators = "mean",
  indic_lags = 1,
  target_lags = 1,
  h = 2
)

forecast(bridge_model)
```

The default mean aggregator is the classic bridge-model setup: each monthly
block is completed first, then averaged to the quarterly frequency before the
target equation is estimated.

The fitted object stores the aligned data that went into estimation and the
future target-period regressor path used for forecasting.

```{r aligned-data}
tail(bridge_model$estimation_set)
bridge_model$forecast_set
```

# Standardized Output

`summary()` and `forecast()` use a stable package-specific layout. The base
output is the same across bridge, MIDAS-style, and direct-alignment
specifications. Additional details, such as optimization summaries or
bootstrap settings, are appended only when they are relevant.

```{r model-summary}
summary(bridge_model)
```

# Direct Alignment

If you set `indic_predict = "direct"`, `bridgr` switches from indicator
forecasting to direct MIDAS-style alignment. In that case, the latest complete
high-frequency blocks are assigned backward to target periods instead of being
forecast forward first.

```{r direct-model}
direct_model <- bridge(
  target = gdp_growth,
  indic = baro,
  indic_predict = "direct",
  indic_aggregators = "unrestricted",
  h = 1
)

forecast(direct_model)
```

This is particularly useful at the ragged edge when you want to work only with
observed high-frequency information and avoid a separate indicator forecasting
step.

# Optional Bootstrap Uncertainty

By default, `bridgr` returns point forecasts only. If you want uncertainty
output, request it at estimation time with `se = TRUE` and a bootstrap
configuration.

```{r bootstrap-example, eval = FALSE}
boot_model <- bridge(
  target = gdp_growth,
  indic = baro,
  indic_predict = "auto.arima",
  indic_aggregators = "mean",
  target_lags = 1,
  h = 1,
  se = TRUE,
  bootstrap = list(type = "block", N = 100, block_length = NULL)
)

forecast(boot_model)
summary(boot_model)
```

The current uncertainty implementation is a conditional block bootstrap on the
final target-frequency estimation sample. It does not re-estimate indicator
forecast models or parametric aggregation weights inside each bootstrap draw.

# Where to Go Next

The vignette
`vignette("mixed-frequency-modeling", package = "bridgr")` compares the main
aggregation strategies and shows how `bridgr` moves from classic bridge models
to unrestricted and parametric MIDAS-style specifications.

The vignette
`vignette("ragged-edge-nowcasting", package = "bridgr")` focuses on
`indic_predict` and the different ways to handle incomplete high-frequency
data at the forecast origin.

The vignette
`vignette("uncertainty-and-scenarios", package = "bridgr")` shows how to work
with conditional block-bootstrap uncertainty and scenario forecasts based on
custom future regressor paths.
