---
title: "Aggregation and Mixed-Frequency Modeling in bridgr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Aggregation and Mixed-Frequency Modeling in bridgr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 6,
  fig.height = 3.5
)
set.seed(2024)
library(bridgr)
```

# The Bridge-to-MIDAS Spectrum

The key design choice in `bridgr` is that classic bridge equations,
unrestricted mixed-frequency regressions, and parametric MIDAS-style
models all share the same estimation, forecasting, and summary workflow.
You can move between them without switching to a different API.

`bridgr` supports several ways to map higher-frequency observations into a
lower-frequency target equation. Those choices span a continuum:

- Bridge-style deterministic aggregation: `"mean"`, `"last"`, `"sum"`, or
  a fixed numeric weight vector.
- Unrestricted mixed-frequency regression: `"unrestricted"`.
- Parametric MIDAS-style weighting: `"expalmon"` and `"beta"`.

This vignette uses a simple monthly-to-quarterly simulation so the
differences between these approaches are easy to see.

# A Small Monthly-to-Quarterly Example

```{r simulation}
n_quarters <- 40
quarter_index <- rep(seq_len(n_quarters), each = 3)
slot <- rep(1:3, times = n_quarters)
monthly_time <- seq(
  as.Date("2010-01-01"),
  by = "month",
  length.out = n_quarters * 3
)
monthly_indicator <- dplyr::tibble(
  time = monthly_time,
  value = 15 + quarter_index * 0.35 +
    ifelse(slot == 1, 0.8 * sin(quarter_index / 2), 0) +
    ifelse(slot == 2, -0.6 * cos(quarter_index / 3), 0) +
    ifelse(slot == 3, 0.7 * sin(quarter_index / 4 + 0.3), 0)
)

quarter_time <- monthly_time[seq(1, length(monthly_time), by = 3)]
quarter_target <- dplyr::tibble(
  time = quarter_time,
  value = 0.5 +
    vapply(
      seq_along(quarter_time),
      function(i) {
        block <- monthly_indicator$value[((i - 1) * 3 + 1):(i * 3)]
        0.2 * block[[1]] + 0.6 * block[[2]] + 0.2 * block[[3]]
      },
      numeric(1)
    ) +
    rep(c(0.1, -0.05, 0.08, -0.02), length.out = length(quarter_time))
)
```

The data-generating process is intentionally driven more by the middle
month of each quarter — the within-quarter weights are `(0.2, 0.6, 0.2)` —
so we can see how the different aggregation schemes react. The monthly
indicator also has slot-specific within-quarter movements, so the
unrestricted specification can estimate three distinct monthly
coefficients.

# Deterministic Bridge Aggregation

```{r deterministic}
mean_model <- mf_model(
  target = quarter_target,
  indic = monthly_indicator,
  indic_predict = "last",
  indic_aggregators = "mean",
  h = 1
)

last_model <- mf_model(
  target = quarter_target,
  indic = monthly_indicator,
  indic_predict = "last",
  indic_aggregators = "last",
  h = 1
)

summary(mean_model)
```

These are classic bridge-model choices. They are easy to interpret and
often work well when you want a transparent rule for within-period
aggregation.

## Fixed numeric weights

If you already have a prior about the within-period shape, pass a numeric
weight vector in a `list()` instead. Weights must have the right length
for the inferred target-period block size and must sum to one.

```{r numeric-weights}
fixed_model <- mf_model(
  target = quarter_target,
  indic = monthly_indicator,
  indic_predict = "last",
  indic_aggregators = list(c(0.1, 0.7, 0.2)),
  h = 1
)

stats::coef(fixed_model)
```

# Unrestricted Mixed-Frequency Regression

```{r unrestricted}
unrestricted_model <- mf_model(
  target = quarter_target,
  indic = monthly_indicator,
  indic_predict = "last",
  indic_aggregators = "unrestricted",
  h = 1
)

stats::coef(unrestricted_model)
```

`"unrestricted"` estimates one coefficient for each within-quarter monthly
observation. In a monthly-on-quarterly example this means three separate
regressors. Notice that the recovered coefficients are close to the
true `(0.2, 0.6, 0.2)` weights used in the simulated DGP — the second
slot dominates, as designed.

This example keeps `indic_predict = "last"` so the comparison isolates
the aggregation choice. The ragged-edge vignette covers
`indic_predict = "direct"`, which skips indicator forecasting and instead
works from the latest observed complete high-frequency blocks.

# Parametric MIDAS-Style Weighting

`bridgr` also supports parametric weighting rules that estimate the
within-period shape from the data while keeping the number of free
parameters small. The `solver_options` list controls the optimizer —
see `?mf_model` for the full set of available controls.

```{r parametric}
expalmon_model <- mf_model(
  target = quarter_target,
  indic = monthly_indicator,
  indic_predict = "last",
  indic_aggregators = "expalmon",
  solver_options = list(seed = 123, n_starts = 1, maxiter = 100),
  h = 1
)

beta_model <- mf_model(
  target = quarter_target,
  indic = monthly_indicator,
  indic_predict = "last",
  indic_aggregators = "beta",
  solver_options = list(
    seed = 123,
    n_starts = 1,
    maxiter = 100,
    start_values = c(2, 2)
  ),
  h = 1
)
```

The fitted object stores both the estimated weight profile and the
underlying parametric coefficients.

```{r weight-comparison}
indicator_id <- expalmon_model$indic_name[[1]]

equal_weights <- rep(1 / 3, 3)
last_weights <- c(0, 0, 1)
true_weights <- c(0.2, 0.6, 0.2)

weights_df <- dplyr::bind_rows(
  dplyr::tibble(model = "mean",        month = 1:3, weight = equal_weights),
  dplyr::tibble(model = "last",        month = 1:3, weight = last_weights),
  dplyr::tibble(model = "true DGP",    month = 1:3, weight = true_weights),
  dplyr::tibble(model = "expalmon",    month = 1:3,
                weight = expalmon_model$parametric_weights[[indicator_id]]),
  dplyr::tibble(model = "beta",        month = 1:3,
                weight = beta_model$parametric_weights[[indicator_id]])
)

ggplot2::ggplot(
  weights_df,
  ggplot2::aes(x = .data$month, y = .data$weight, color = .data$model)
) +
  ggplot2::geom_line(linewidth = 0.8) +
  ggplot2::geom_point(size = 2) +
  ggplot2::scale_x_continuous(breaks = 1:3) +
  ggplot2::labs(
    title = "Within-quarter weight profiles",
    x = "Month within quarter",
    y = "Weight"
  ) +
  theme_bridgr()
```

The parametric profiles concentrate mass on the middle month, matching
the simulated DGP. The fixed `"mean"` rule spreads mass evenly across
months and `"last"` puts all of it on the final month.

# Forecast Comparison

All of these models share the same downstream interface.

```{r forecast-comparison}
forecasts_df <- dplyr::bind_rows(
  dplyr::tibble(model = "mean",
                forecast = as.numeric(forecast(mean_model)$mean)),
  dplyr::tibble(model = "last",
                forecast = as.numeric(forecast(last_model)$mean)),
  dplyr::tibble(model = "unrestricted",
                forecast = as.numeric(forecast(unrestricted_model)$mean)),
  dplyr::tibble(model = "expalmon",
                forecast = as.numeric(forecast(expalmon_model)$mean)),
  dplyr::tibble(model = "beta",
                forecast = as.numeric(forecast(beta_model)$mean))
)

ggplot2::ggplot(
  forecasts_df,
  ggplot2::aes(x = .data$model, y = .data$forecast, fill = .data$model)
) +
  ggplot2::geom_col(width = 0.6, show.legend = FALSE) +
  ggplot2::labs(
    title = "One-step-ahead forecast by aggregation strategy",
    x = NULL, y = "Forecast"
  ) +
  theme_bridgr()
```

# Picking an Aggregator on Out-of-Sample Performance

A rough rule-of-thumb guide to aggregator choice is given at the end of
this vignette. In practice the best way to choose is to evaluate
candidate models on out-of-sample forecasts. The example below uses a
short pseudo-real-time rolling-origin scheme on the simulated data, with
one-quarter-ahead forecasts.

```{r evaluation, message = FALSE, warning = FALSE}
origins <- (n_quarters - 13):(n_quarters - 4)
methods <- c("mean", "unrestricted", "expalmon")

eval_rows <- list()
for (m in methods) {
  for (origin in origins) {
    train <- dplyr::slice(quarter_target, seq_len(origin))
    truth <- quarter_target$value[[origin + 1]]
    fit <- mf_model(
      target = train,
      indic = monthly_indicator,
      indic_predict = "last",
      indic_aggregators = m,
      h = 1,
      solver_options = if (m == "expalmon") {
        list(seed = 1, n_starts = 1, maxiter = 100)
      } else {
        NULL
      }
    )
    fc <- as.numeric(forecast(fit)$mean)[[1]]
    eval_rows[[length(eval_rows) + 1]] <- dplyr::tibble(
      method = m, origin = origin, forecast = fc, actual = truth
    )
  }
}

eval_df <- dplyr::bind_rows(eval_rows)
eval_summary <- eval_df |>
  dplyr::group_by(.data$method) |>
  dplyr::summarise(
    rmse = sqrt(mean((.data$forecast - .data$actual)^2)),
    mae = mean(abs(.data$forecast - .data$actual)),
    .groups = "drop"
  )

eval_summary
```

Even on this small example the unrestricted and parametric specifications
recover the within-quarter shape and outperform the deterministic mean.
For a real application you would typically use a longer evaluation
window, multiple horizons, and richer scoring rules.

# Choosing an Aggregation Strategy

As a rough guide:

- Use deterministic bridge aggregation like `"mean"`, `"last"`, or
  `"sum"` when you want a transparent and stable nowcasting rule.
- Use `"unrestricted"` when the frequency gap is small and you want each
  within-period observation to have its own coefficient.
- Use `"expalmon"` or `"beta"` when you want data-driven within-period
  weights but would like a more parsimonious parameterization than
  `"unrestricted"`.
- Use `indic_predict = "direct"` when you want direct alignment based
  only on the latest observed complete high-frequency blocks.

## Related packages

If you have used [`midasr`](https://CRAN.R-project.org/package=midasr),
the closest `bridgr` analogues to its MIDAS regressors are
`indic_aggregators = "expalmon"` and `indic_aggregators = "beta"`. The
`"unrestricted"` aggregator plays a role similar to `midasr`'s U-MIDAS
specification. Compared with
[`midasml`](https://CRAN.R-project.org/package=midasml), which targets
high-dimensional regularized mixed-frequency models, `bridgr` is aimed at
lower-dimensional applied forecasting workflows.
