---
title: "Aggregation and Mixed-Frequency Modeling in bridgr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Aggregation and Mixed-Frequency Modeling in bridgr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

library(bridgr)
```

# The Bridge-to-MIDAS Spectrum

`bridgr` supports several ways to map higher-frequency observations into a
lower-frequency target equation. Those choices span a continuum:

- Bridge-style deterministic aggregation: `"mean"`, `"last"`, `"sum"`, or
  fixed numeric weights.
- Unrestricted mixed-frequency regression: `"unrestricted"`.
- Parametric MIDAS-style weighting: `"expalmon"`, `"beta"`, and `"legendre"`.

This vignette uses a simple monthly-to-quarterly simulation so the differences
between these approaches are easy to see.

# A Small Monthly-to-Quarterly Example

```{r simulation}
n_quarters <- 40
quarter_index <- rep(seq_len(n_quarters), each = 3)
slot <- rep(1:3, times = n_quarters)
monthly_time <- seq(
  as.Date("2010-01-01"),
  by = "month",
  length.out = n_quarters * 3
)
monthly_indicator <- dplyr::tibble(
  time = monthly_time,
  value = 15 + quarter_index * 0.35 +
    ifelse(slot == 1, 0.8 * sin(quarter_index / 2), 0) +
    ifelse(slot == 2, -0.6 * cos(quarter_index / 3), 0) +
    ifelse(slot == 3, 0.7 * sin(quarter_index / 4 + 0.3), 0)
)

quarter_time <- monthly_time[seq(1, length(monthly_time), by = 3)]
quarter_target <- dplyr::tibble(
  time = quarter_time,
  value = 0.5 +
    vapply(
      seq_along(quarter_time),
      function(i) {
        block <- monthly_indicator$value[((i - 1) * 3 + 1):(i * 3)]
        0.2 * block[[1]] + 0.6 * block[[2]] + 0.2 * block[[3]]
      },
      numeric(1)
    ) +
    rep(c(0.1, -0.05, 0.08, -0.02), length.out = length(quarter_time))
)
```

The target is intentionally driven more by the middle month of each quarter, so
we can see how the different aggregation schemes react. The monthly indicator
also has slot-specific within-quarter movements, so the unrestricted
specification can estimate three distinct monthly coefficients.

# Deterministic Bridge Aggregation

```{r deterministic}
mean_model <- bridge(
  target = quarter_target,
  indic = monthly_indicator,
  indic_predict = "last",
  indic_aggregators = "mean",
  h = 1
)

last_model <- bridge(
  target = quarter_target,
  indic = monthly_indicator,
  indic_predict = "last",
  indic_aggregators = "last",
  h = 1
)

summary(mean_model)
summary(last_model)
```

These are classic bridge-model choices. They are easy to interpret and often
work well when you want a transparent rule for within-period aggregation.

# Unrestricted Mixed-Frequency Regression

```{r unrestricted}
unrestricted_model <- bridge(
  target = quarter_target,
  indic = monthly_indicator,
  indic_predict = "last",
  indic_aggregators = "unrestricted",
  h = 1
)

summary(unrestricted_model)
```

`"unrestricted"` estimates one coefficient for each within-quarter monthly
observation. In a monthly-on-quarterly example this means three separate
regressors. This example keeps `indic_predict = "last"` so the comparison
isolates the aggregation choice; combining `"unrestricted"` with
`indic_predict = "direct"` gives a direct MIDAS-style alignment and is covered
in the ragged-edge vignette.

# Parametric MIDAS-Style Weighting

`bridgr` also supports parametric weighting rules that estimate the
within-period shape from the data while keeping the number of free parameters
small.

```{r parametric}
expalmon_model <- bridge(
  target = quarter_target,
  indic = monthly_indicator,
  indic_predict = "last",
  indic_aggregators = "expalmon",
  solver_options = list(seed = 123, n_starts = 1, maxiter = 100),
  h = 1
)

beta_model <- bridge(
  target = quarter_target,
  indic = monthly_indicator,
  indic_predict = "last",
  indic_aggregators = "beta",
  solver_options = list(
    seed = 123,
    n_starts = 1,
    maxiter = 100,
    start_values = c(2, 2)
  ),
  h = 1
)

legendre_model <- bridge(
  target = quarter_target,
  indic = monthly_indicator,
  indic_predict = "last",
  indic_aggregators = "legendre",
  solver_options = list(seed = 123, n_starts = 1, maxiter = 100),
  h = 1
)
```

The fitted object stores both the estimated weight profile and the underlying
parametric coefficients.

```{r weight-comparison}
dplyr::bind_rows(
  dplyr::tibble(
    model = "expalmon",
    month = seq_along(expalmon_model$parametric_weights$indic),
    weight = expalmon_model$parametric_weights$indic
  ),
  dplyr::tibble(
    model = "beta",
    month = seq_along(beta_model$parametric_weights$indic),
    weight = beta_model$parametric_weights$indic
  ),
  dplyr::tibble(
    model = "legendre",
    month = seq_along(legendre_model$parametric_weights$indic),
    weight = legendre_model$parametric_weights$indic
  )
)
```

# Forecast Comparison

All of these models share the same downstream interface.

```{r forecast-comparison}
dplyr::bind_rows(
  dplyr::tibble(
    model = "mean",
    forecast = as.numeric(forecast(mean_model)$mean)
  ),
  dplyr::tibble(
    model = "last",
    forecast = as.numeric(forecast(last_model)$mean)
  ),
  dplyr::tibble(
    model = "unrestricted",
    forecast = as.numeric(forecast(unrestricted_model)$mean)
  ),
  dplyr::tibble(
    model = "expalmon",
    forecast = as.numeric(forecast(expalmon_model)$mean)
  ),
  dplyr::tibble(
    model = "beta",
    forecast = as.numeric(forecast(beta_model)$mean)
  ),
  dplyr::tibble(
    model = "legendre",
    forecast = as.numeric(forecast(legendre_model)$mean)
  )
)
```

# Choosing an Aggregation Strategy

As a rough guide:

- Use deterministic bridge aggregation when you want a transparent and stable
  nowcasting rule.
- Use `"unrestricted"` when the frequency gap is small and you want each
  within-period observation to have its own coefficient.
- Use `"expalmon"`, `"beta"`, or `"legendre"` when you want data-driven
  within-period weights but would like a more parsimonious parameterization
  than `"unrestricted"`.
- Use `indic_predict = "direct"` when you want direct MIDAS-style alignment
  based only on the latest observed high-frequency blocks.

The key design choice in `bridgr` is that all of these specifications share
the same estimation, forecasting, and summary workflow. You can therefore move
between bridge and MIDAS-style models without switching to a different API.
