---
title: "Ragged-Edge Nowcasting with bridgr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Ragged-Edge Nowcasting with bridgr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

library(bridgr)
```

# Why Ragged Edges Matter

Mixed-frequency nowcasting usually happens at the ragged edge: the
lower-frequency target has not been released yet, and the higher-frequency
indicators are only partially observed for the current target period.

`bridgr` handles that situation through `indic_predict`, which controls how
the missing high-frequency tail is completed before the target equation is
estimated or forecast.

The currently supported options are:

- `"last"`: extend the latest available high-frequency observation.
- `"mean"`: extend the mean of the latest available high-frequency block.
- `"auto.arima"`: fit an automatic ARIMA model to each indicator and forecast
  the missing observations.
- `"ets"`: fit an ETS model to each indicator and forecast the missing
  observations.
- `"direct"`: do not forecast the indicator at all. Instead, align the latest
  observed complete high-frequency blocks directly to the target periods and
  average them within each target period.

# A Ragged-Edge Example

We start from the package's quarterly GDP growth series and monthly barometer,
then remove the last quarterly GDP observation and the last two monthly
indicator observations. That creates a clean nowcast setup: the final quarter
is now forecasted rather than estimated, and the monthly indicator is only
partially observed for that forecast quarter.

```{r data}
gdp_growth <- suppressMessages(tsbox::ts_na_omit(tsbox::ts_pc(gdp)))
gdp_nowcast <- gdp_growth |>
  dplyr::slice_head(n = nrow(gdp_growth) - 1)

baro_ragged <- baro |>
  dplyr::slice_head(n = nrow(baro) - 2)

tail(gdp_growth)
tail(gdp_nowcast)
tail(baro_ragged)
```

The target history now ends one quarter earlier, while the monthly indicator
contains only a partial block for the quarter we want to nowcast.

# Comparing Indicator Completion Rules

```{r compare-methods}
predict_methods <- c("last", "mean", "auto.arima", "ets")

models <- lapply(
  predict_methods,
  function(method) {
    bridge(
      target = gdp_nowcast,
      indic = baro_ragged,
      indic_predict = method,
      indic_aggregators = "mean",
      target_lags = 1,
      h = 1
    )
  }
)
names(models) <- predict_methods

dplyr::bind_rows(
  lapply(
    predict_methods,
    function(method) {
      fc <- forecast(models[[method]])
      dplyr::tibble(
        method = method,
        forecast_time = fc$time,
        forecast = as.numeric(fc$mean)
      )
    }
  )
)
```

The model specification is the same in each case. The only difference is how
the missing monthly observations are completed before the indicator is
aggregated to the quarterly frequency.

# Looking at the Forecast Regressor Set

The easiest way to see what each method is doing is to inspect
`forecast_set`, which stores the target-period regressors used for the final
forecast.

```{r forecast-sets}
lapply(
  models[c("last", "mean")],
  function(model) {
    model$forecast_set
  }
)
```

With `"last"`, the last observed monthly value is extended forward. With
`"mean"`, the mean of the latest available high-frequency block is extended
forward instead.

# Direct Alignment

`"direct"` takes a different route. The indicator is not forecasted. Instead,
`bridgr` assigns the latest complete high-frequency blocks backward to the
target periods and averages them within each target period.

```{r direct}
direct_model <- bridge(
  target = gdp_nowcast,
  indic = baro_ragged,
  indic_predict = "direct",
  h = 1
)

forecast(direct_model)
```

This is especially useful when you want to avoid a separate indicator
forecasting step and prefer to work only with observed high-frequency data.

# When to Use Which Option

As a rough guide:

- Use `"last"` when a simple carry-forward rule is acceptable.
- Use `"mean"` when you want a stable deterministic fill based on the latest
  high-frequency block.
- Use `"auto.arima"` or `"ets"` when the indicator has meaningful time-series
  structure and a separate extrapolation step is sensible.
- Use `"direct"` when you want direct alignment without indicator
  forecasting and want to rely only on the latest observed complete
  high-frequency blocks.

The good part is that the downstream workflow does not change. The same
`summary()` and `forecast()` interface works across all of these nowcasting
choices.
