---
title: "Bootstrap Uncertainty and Scenario Forecasting"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Bootstrap Uncertainty and Scenario Forecasting}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

library(bridgr)
```

# Overview

`bridgr` can return point forecasts only, or it can also compute bootstrap
uncertainty for the fitted target equation and predictive intervals for
forecasts.

The relevant estimation arguments are:

- `se = FALSE`: point forecasts only.
- `se = TRUE`: compute bootstrap coefficient uncertainty and predictive
  forecast intervals.
- `bootstrap = list(type = "block", N = 100, block_length = NULL)`: control
  the bootstrap.

At the moment, `"block"` is the only supported bootstrap type.

# Fitting a Model with Bootstrap Uncertainty

```{r bootstrap-model}
gdp_growth <- suppressMessages(tsbox::ts_na_omit(tsbox::ts_pc(gdp)))

boot_model <- bridge(
  target = gdp_growth,
  indic = baro,
  indic_predict = "auto.arima",
  indic_aggregators = "mean",
  indic_lags = 1,
  target_lags = 1,
  h = 2,
  se = TRUE,
  bootstrap = list(type = "block", N = 20, block_length = NULL)
)
```

The current implementation uses a conditional block bootstrap on the final
target-frequency estimation sample. It does not re-estimate indicator forecast
models or parametric aggregation weights inside each bootstrap draw, but it
does add simulated future target shocks to the stored forecast draws.

# Forecast Output

Once the model has been estimated with `se = TRUE`, `forecast()` returns a
standardized forecast object with:

- `mean`
- `se`
- `lower`
- `upper`
- `forecast_set`
- bootstrap metadata

```{r forecast-output}
fc <- forecast(boot_model)

fc
fc$bootstrap
```

The intervals are empirical predictive intervals based on the stored bootstrap
forecast draws.

# Summary Output

The same bootstrap run also feeds into `summary()`.

```{r summary-output}
summary(boot_model)
```

The printed summary keeps the same base layout as a point-estimate model and
adds the uncertainty section only when bootstrap output is available.

# Scenario Forecasting with `xreg`

If you want to forecast the target under a different future regressor path,
pass a custom `xreg` object to `forecast()`. The custom regressor names must
match the ones used in the fitted target equation.

```{r scenario}
scenario_xreg <- dplyr::tibble(
  id = rep(boot_model$regressor_names, each = nrow(boot_model$forecast_set)),
  time = rep(
    boot_model$forecast_set$time,
    times = length(boot_model$regressor_names)
  ),
  value = c(
    boot_model$forecast_set$baro + 5,
    boot_model$forecast_set$baro_lag1 + 5
  )
)

scenario_fc <- forecast(boot_model, xreg = scenario_xreg)

dplyr::tibble(
  time = fc$time,
  baseline = as.numeric(fc$mean),
  scenario = as.numeric(scenario_fc$mean)
)
```

When a model contains stored bootstrap draws, `forecast(..., xreg = ...)`
reuses the bootstrap refits and evaluates them on the scenario regressor path.
That keeps the predictive uncertainty handling consistent between the baseline
forecast and scenario forecasts.

# Point Forecasts Only

If you are only interested in point estimates, leave `se = FALSE`. In that
case, `bootstrap` is ignored and `forecast()` still returns the same object
shape, with `NA` uncertainty fields.

```{r point-only}
point_model <- bridge(
  target = gdp_growth,
  indic = baro,
  indic_predict = "auto.arima",
  indic_aggregators = "mean",
  indic_lags = 1,
  target_lags = 1,
  h = 1,
  se = FALSE,
  bootstrap = list(type = "block", N = 20)
)

forecast(point_model)
```

# Interpretation

These intervals are conditional on the aligned target-frequency design matrix.
That means they are predictive intervals for the final target equation given
the regressor path, rather than full end-to-end uncertainty for the whole
mixed-frequency system.
