title: "Uncertainty and Scenario Forecasting"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Uncertainty and Scenario Forecasting}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

library(bridgr)
```

# Overview

`bridgr` can return point forecasts only, or it can also compute coefficient
uncertainty for the fitted target equation and prediction intervals for
forecasts.

The relevant estimation arguments are:

- `se = FALSE`: point forecasts only.
- `se = TRUE`: compute HAC or Delta-HAC coefficient uncertainty and
  prediction intervals.
- `bootstrap = list(N = 100, block_length = NULL)`: control the number of
  predictive simulation paths. `block_length` is only used when
  `full_system_bootstrap = TRUE`.
- `full_system_bootstrap = TRUE`: replace the default residual-resampling
  prediction intervals with a full-system target-period block bootstrap.
  Because this refits the full bridge workflow on every draw, it can be
  substantially slower.

# Fitting a Model with Default Uncertainty

```{r bootstrap-model}
gdp_growth <- suppressMessages(tsbox::ts_na_omit(tsbox::ts_pc(gdp)))

boot_model <- bridge(
  target = gdp_growth,
  indic = baro,
  indic_predict = "auto.arima",
  indic_aggregators = "mean",
  indic_lags = 1,
  target_lags = 1,
  h = 2,
  se = TRUE,
  bootstrap = list(N = 20, block_length = NULL)
)
```

`bridgr` computes HAC standard errors for the linear bridge equation, or
Delta-HAC standard errors when parametric aggregation weights are estimated
jointly. Forecast uncertainty is obtained by simulating from resampled
centered residuals of the fitted target equation.

# Forecast Output

Once the model has been estimated with `se = TRUE`, `forecast()` returns a
standardized forecast object with:

- `mean`
- `se`
- `lower`
- `upper`
- `forecast_set`
- uncertainty metadata

```{r forecast-output}
fc <- forecast(boot_model)

fc
fc$bootstrap
```

The intervals are empirical prediction intervals based on the stored residual-
resampling forecast draws.

# Summary Output

The same uncertainty configuration also feeds into `summary()`.

```{r summary-output}
summary(boot_model)
```

The printed summary keeps the same base layout as a point-estimate model and
adds the uncertainty section only when uncertainty output is available.

# Scenario Forecasting with `xreg`

If you want to forecast the target under a different future regressor path,
pass a custom `xreg` object to `forecast()`. The custom regressor names must
match the ones used in the fitted target equation.

```{r scenario}
scenario_xreg <- dplyr::tibble(
  id = rep(boot_model$xreg_names, each = nrow(boot_model$forecast_base_set)),
  time = rep(
    boot_model$forecast_base_set$time,
    times = length(boot_model$xreg_names)
  ),
  value = c(
    boot_model$forecast_base_set$baro + 5,
    boot_model$forecast_base_set$baro_lag1 + 5
  )
)

scenario_fc <- forecast(boot_model, xreg = scenario_xreg)

dplyr::tibble(
  time = fc$time,
  baseline = as.numeric(fc$mean),
  scenario = as.numeric(scenario_fc$mean)
)
```

Scenario forecasts reuse the same uncertainty method and evaluate it on the
supplied regressor path.

# Optional Full-System Bootstrap

If you want to propagate uncertainty through the full bridge workflow,
including indicator completion and aggregation, set
`full_system_bootstrap = TRUE`. This can be substantially slower than the
default residual-resampling intervals because each bootstrap draw re-estimates
the full bridge pipeline.

```{r full-system-bootstrap}
full_model <- bridge(
  target = gdp_growth,
  indic = baro,
  indic_predict = "auto.arima",
  indic_aggregators = "mean",
  indic_lags = 1,
  target_lags = 1,
  h = 2,
  se = TRUE,
  full_system_bootstrap = TRUE,
  bootstrap = list(N = 20, block_length = NULL)
)

forecast(full_model)$bootstrap
```

# Point Forecasts Only

If you are only interested in point estimates, leave `se = FALSE`. In that
case, `bootstrap` is ignored and `forecast()` still returns the same object
shape, with `NA` uncertainty fields.

```{r point-only}
point_model <- bridge(
  target = gdp_growth,
  indic = baro,
  indic_predict = "auto.arima",
  indic_aggregators = "mean",
  indic_lags = 1,
  target_lags = 1,
  h = 1,
  se = FALSE,
  bootstrap = list(N = 20)
)

forecast(point_model)
```

# Interpretation

With the default `se = TRUE`, these intervals are residual-resampling
prediction intervals for the fitted target equation. When
`full_system_bootstrap = TRUE`, the intervals instead come from a full-system
block bootstrap that resamples target-period blocks from the aligned mixed-
frequency system, refits the bridge workflow on each draw, and evaluates the
resulting future paths.
