---
title: "Uncertainty and Scenario Forecasting"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Uncertainty and Scenario Forecasting}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 6,
  fig.height = 3.5
)
set.seed(2024)
library(bridgr)
```

# Overview

`bridgr` can return point forecasts only, or it can also compute
coefficient uncertainty for the fitted target equation and prediction
intervals for forecasts.

The relevant estimation arguments are:

- `se = FALSE`: point forecasts only.
- `se = TRUE`: compute HAC or Delta-HAC coefficient uncertainty and
  prediction intervals.
- `bootstrap = list(N = 100, block_length = NULL)`: control the number
  of predictive simulation paths. `block_length` is only used when
  `full_system_bootstrap = TRUE`, where it sets the size of the
  contiguous target-period blocks resampled in each bootstrap draw —
  larger blocks preserve more temporal dependence.
- `full_system_bootstrap = TRUE`: replace the default
  residual-resampling prediction intervals and HAC / Delta-HAC
  coefficient standard errors with a full-system target-period block
  bootstrap. Because this refits the full mixed-frequency workflow on
  every draw, it can be substantially slower.

# Fitting a Model with Default Uncertainty

```{r bootstrap-model}
gdp_growth <- suppressMessages(tsbox::ts_na_omit(tsbox::ts_pc(gdp)))

boot_model <- mf_model(
  target = gdp_growth,
  indic = baro,
  indic_predict = "auto.arima",
  indic_aggregators = "mean",
  indic_lags = 1,
  target_lags = 1,
  h = 2,
  se = TRUE,
  bootstrap = list(N = 20, block_length = NULL)
)
```

`bridgr` computes HAC standard errors for the linear target equation, or
Delta-HAC standard errors when parametric aggregation weights are
estimated jointly. Forecast uncertainty is obtained by simulating from
resampled centered residuals of the fitted target equation.

# Forecast Output

Once the model has been estimated with `se = TRUE`, `forecast()` returns
a standardized forecast object with:

- `mean`
- `se`
- `lower`
- `upper`
- `forecast_set`
- uncertainty metadata

```{r forecast-output}
fc <- forecast(boot_model)

fc
fc$bootstrap
```

The intervals are empirical prediction intervals based on the stored
residual-resampling forecast draws.

# Reporting Only Forecasts Below a Width Tolerance

In practice prediction intervals widen with horizon. A common reporting
choice is to publish only forecast horizons whose interval is narrower
than some application-specific tolerance — for example, only publish
when the 95% half-width is below a chosen threshold. The example below
applies that rule with a tolerance of `1.5` percentage points.

```{r width-tolerance}
forecast_table <- dplyr::tibble(
  time = fc$time,
  mean = as.numeric(fc$mean),
  lower_95 = fc$lower[, "95%"],
  upper_95 = fc$upper[, "95%"]
) |>
  dplyr::mutate(
    half_width_95 = (.data$upper_95 - .data$lower_95) / 2
  )

tolerance <- 1.5
dplyr::filter(forecast_table, .data$half_width_95 <= tolerance)
```

Tighten or relax `tolerance` depending on how much uncertainty is
acceptable in the use case.

# Summary Output

The same uncertainty configuration also feeds into `summary()`.

```{r summary-output}
summary(boot_model)
```

The printed summary keeps the same base layout as a point-estimate model
and adds the uncertainty section only when uncertainty output is
available.

# Scenario Forecasting with `xreg`

If you want to forecast the target under a different future regressor
path, pass a custom `xreg` object to `forecast()`. The custom regressor
names must match the ones used in the fitted target equation.

The example below defines a *baseline* scenario (the indicator continues
on its model-implied path) and a *negative shock* scenario (the
indicator drops by 5 points relative to baseline). Both scenarios share
the fitted target equation and the same uncertainty method.

```{r scenarios}
make_xreg <- function(level_shift) {
  dplyr::tibble(
    id = rep(boot_model$xreg_names, each = nrow(boot_model$forecast_base_set)),
    time = rep(
      boot_model$forecast_base_set$time,
      times = length(boot_model$xreg_names)
    ),
    value = c(
      boot_model$forecast_base_set$baro + level_shift,
      boot_model$forecast_base_set$baro_lag1 + level_shift
    )
  )
}

fc_baseline <- forecast(boot_model, xreg = make_xreg(0))
fc_shock    <- forecast(boot_model, xreg = make_xreg(-5))

scenario_df <- dplyr::bind_rows(
  dplyr::tibble(
    scenario = "baseline",
    time = fc_baseline$time,
    mean = as.numeric(fc_baseline$mean),
    lower = fc_baseline$lower[, "95%"],
    upper = fc_baseline$upper[, "95%"]
  ),
  dplyr::tibble(
    scenario = "shock (-5)",
    time = fc_shock$time,
    mean = as.numeric(fc_shock$mean),
    lower = fc_shock$lower[, "95%"],
    upper = fc_shock$upper[, "95%"]
  )
)

ggplot2::ggplot(
  scenario_df,
  ggplot2::aes(x = .data$time, color = .data$scenario, fill = .data$scenario)
) +
  ggplot2::geom_ribbon(
    ggplot2::aes(ymin = .data$lower, ymax = .data$upper),
    alpha = 0.2, color = NA
  ) +
  ggplot2::geom_line(ggplot2::aes(y = .data$mean), linewidth = 0.8) +
  ggplot2::geom_point(ggplot2::aes(y = .data$mean), size = 2) +
  ggplot2::labs(
    title = "Forecast under baseline and shock scenarios",
    x = NULL, y = "GDP growth forecast"
  ) +
  theme_bridgr()
```

Scenario forecasts reuse the same uncertainty method and evaluate it on
the supplied regressor path.

# Optional Full-System Bootstrap

If you want to propagate uncertainty through the full mixed-frequency
workflow, including indicator completion and aggregation, set
`full_system_bootstrap = TRUE`. This can be substantially slower than
the default residual-resampling intervals because each bootstrap draw
re-estimates the full pipeline. In that mode, both the reported
coefficient standard errors and the forecast intervals are based on the
bootstrap draws.

```{r full-system-bootstrap}
full_model <- mf_model(
  target = gdp_growth,
  indic = baro,
  indic_predict = "auto.arima",
  indic_aggregators = "mean",
  indic_lags = 1,
  target_lags = 1,
  h = 2,
  se = TRUE,
  full_system_bootstrap = TRUE,
  bootstrap = list(N = 20, block_length = NULL)
)

forecast(full_model)$bootstrap
```

# Point Forecasts Only

If you are only interested in point estimates, leave `se = FALSE`. In
that case, `bootstrap` is ignored and `forecast()` still returns the
same object shape, with `NA` uncertainty fields.

```{r point-only}
point_model <- mf_model(
  target = gdp_growth,
  indic = baro,
  indic_predict = "auto.arima",
  indic_aggregators = "mean",
  indic_lags = 1,
  target_lags = 1,
  h = 1,
  se = FALSE,
  bootstrap = list(N = 20)
)

forecast(point_model)
```

# Interpretation

With the default `se = TRUE`, intervals are residual-resampling
prediction intervals for the fitted target equation, and coefficient
standard errors come from HAC or Delta-HAC estimation. When
`full_system_bootstrap = TRUE`, both the coefficient standard errors and
the forecast intervals come from a full-system block bootstrap that
resamples target-period blocks from the aligned mixed-frequency system,
refits the workflow on each draw, and evaluates the resulting future
paths.
